---
title       : "Performance of students in Math and Portugues"
shorttitle    : ""
author: 
- name     : "Jack Sparrow"
  affiliation  : "1"
  corresponding : yes  # Define only one corresponding author
  address    : "Missionsstrasse 62A 4053 Basel Switzerland"
  email     : "sparrow@pmail.com"
  role:     # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
   - Conceptualization
   - Writing - Original Draft Preparation
   - Writing - Review & Editing
   - Resources
   - Validation 
   - Visualization 
   - Project Administration 
   - Formal Analysis
- name     : "Other Folks as Per Order on Doc"
  affiliation  : "2"
  role:
   - Writing - Review & Editing
   - Data Curation
   - Resources
affiliation:
- id      : "1"
  institution  : "Harrisburg University of Science and Technology"
- id      : "2"
  institution  : "Other Instituions"
authornote: |
  
abstract: | 
  This data approach student achievement in secondary education of two Portuguese schools.The data attributes include student grades, demographic, social and school related features)and it was collected by using school reports and questionnaires. Two datasets are providedregarding the performance in two distinct subjects Mathematics (mat) and Portugueselanguage (por)
 
keywords     : "apa, R, markdown"
floatsintext   : no
figurelist    : no
tablelist     : no
footnotelist   : no
linenumbers    : yes
mask       : no
draft       : no
classoption    : "man"
bibliography: r-references.bib
header-includes:
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}
    
    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother

csl               : "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass     : "apa7"
output            : papaja::apa6_pdf
---

```{r include = FALSE, eval=T}

# Set default chunk options (can be overridden in later chunks)
knitr::opts_chunk$set(echo = FALSE, 
                      eval = TRUE,
                      message = FALSE, 
                      fig.width = 4, 
                      fig.height = 4,
                      fig.align = 'center',
                      out.width = "100%")

# Load packages
library("papaja")
library("knitr")
library("tidyverse")
library("ggcorrplot") # install if needed


```


```{r, eval=F}
# read in data
student.math <-

student.por <-

```


What is the relationship between student performance in language and mathematics tasks? This is an important question that has been studied extensively. For example, @horwitz1986foreign found that students frequently feel anxiety in foreign language classes. XXX combined several studies on language achievement and found that language-minority students may need special treatment plans. Interestingly, language appears to be related to performance in mathematics [XXX]. In one study based on a survey of 1,174 8th grade students, XXX found that students who were English language learners (ELLs) scored lower on math tests than proficient speakers of English.

The purpose of the present research was to see if previous results replicate in a new sample of language and mathematics learners. To test this, we analysed data of student performance in Mathematics and Portuguese classes.

# Methods

## Participants

Data were collected from the UCI machine learning repository at <http://archive.ics.uci.edu/ml/datasets/Student+Performance>.  Data from `r  1+1` students in a Mathematics class, and `r 1+1` students in a Portuguese class were collected.

## Procedure

The primary measures were three exam scores taken at the beginning, middle, and end of each class. 

# Results

All analyses were conducted using R [@R] and compiled using the papaja package [@aust2015papaja].

Distributions of the three exam scores for the Mathematics and Portuguese classes are presented in Figure 1. Correlations between numeric predictors in the Math data are shown in Figure 2.

```{r summary, eval=F}
# Create long version of the Portuguese data
student_por_long <- student.por %>% 
  select(_, _, _) %>% 
  # pivot_longer takes columns and makes them into rows. This is useful for plotting 
  pivot_longer(_:_,names_to="Exam", values_to= "score") %>%
  mutate(class = "Portuguese")

# Create long version of the math data
student_math_long <- student.math %>% 
 select(_,_,_) %>% 
  pivot_longer(_:_,names_to="Exam", values_to= "score") %>%
  mutate(class = "Math")


```


```{r fig1, fig.width = 6, fig.height = 4, fig.cap= "XXX", eval = F}
# Figure 1: Histograms of distributions of exam scores

# Combine two long versions
student_all <- rbind(student.por.long, student.math.long)

# Create grid of histograms
ggplot(data = student_all, 
       aes(x = score)) +
  facet_wrap(~ class + Exam) +
  geom_histogram(col = "black", fill = "white", bins = 15) +
  scale_x_continuous(limits = c(0, 20)) +
  theme_minimal()
```


```{r fig2, fig.width = 6, fig.height = 6, fig.cap= "XXX", eval = F, out.width = "70%"}
# Figure 2: Correlation plot

# Get matrix of correlations
math_cor <- cor(student.por[c("studytime", "failures", "freetime", "goout", "absences", "g1", "g2", "g3")])

# Matrix of p values
math_p <- ggcorrplot::cor_pmat(student.math[c("studytime", "failures", "freetime", "goout", "absences", "g1", "g2", "g3")])

# Create correlation plot with ggpcorrplot
ggcorrplot::ggcorrplot(math_cor, 
                       method = "circle", 
                       lab = TRUE,        # Include correlation coefficient labels
                       lab_size = 3,      # Reduce label size a bit
                       p.mat = math_p)    # Include p-values so non-significant values have X's

```

Descriptive statistics of grades separated by sex and school are presented in Tables 1 and 2. Grades tended to increase over the course of the semester. For example, the mean grade in the first Portuguese exam was `r 1+1` which increased to `r 1+1` by the last exam.

```{r tbl1, results='asis', eval = F}
# Create summary table of Portuguese data
Portuguese.tbl <- student.por %>% 
  group_by(sex, school) %>%
  summarise(
    "Exam 1" = round(mean(g1), 2),
    "Exam 2" = round(mean(g2), 2),
    "Exam 3" = round(mean(g3), 2)
  )

# Print the table!
# also be sure to include the results='asis' chunk option!
papaja::apa_table(Portuguese.tbl, 
                  caption = "XXX")
```

```{r tbl2, results='asis', eval = F}
# Create summary table of Math data
math.tbl <- student.math %>% group_by(sex, school) %>%
  summarise(
    "Exam 1" = round(mean(g1), 2),
    "Exam 2" = round(mean(g2), 2),
    "Exam 3" = round(mean(g3), 2)
  )

#creates tables in apa format
papaja::apa_table(math.tbl, 
                  caption = "XXX")
```

```{r ttests, eval = F}
# regression  of Exam 1 Portuguese grades between sexes
sex.por.lm <- lm(g1 ~ sex, data = student.por)

# lm of Exam 1 math grades between sexes


```

Did men and women perform differently on the first exams in each class? To test this, we conducted a linear regression  on first exam scores as a function of sex. The lm on Portuguese exam 1 was significant `r 1+1`, showing that women performed better than men on the first Portuguese exam. The lm on Math exam 1 was non-significant `r 1+1`, showing no evidence for a difference between men and women on Math exam 1.


```{r, results='asis', eval = F}
g1_por_Anova <- car::Anova(mod = lm(g1 ~ school + sex + guardian,
                                    data = student.por),
                           type = "III")

papaja::apa_table(apa_print(g1_por_Anova)$table, caption = "ANOVA on period 1 Portuguese scores.")
```


# Discussion

Understanding the relationship between language and math performance is important for understanding learning. Our results are generally in line with @abedi2001language who found a relationship between language and mathematics performance.


# References



